\texttt{jaxoplanet}: Hardware-Accelerated Computation of Stellar Light Curves

### 2024-09-23 latex/paper.tex
feat: bits on the limb-darkened case.

The surface of a spherical body can be naturally described by spherical harmonics: a set of special functions that can be used as orthonormal basis to represent any spherical surface map. If $\bvec{\tilde y}$ denotes the spherical harmonics basis of degree $l_{max}$ (defined in \citealt[section X]{starry}) and $\bvec{y}$ a vector that contains the coefficients describing the surface in this basis, the intensity $I$ of the map projected onto the sky-projected $(x, y)$ plane of the observer is
$$I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{y}.$$
In order to orient this surface, for example with a given inclination, obliquity and phase around its rotation axis, we define a rotation matrix $\bvec{R}$ that acts on the spherical harmonics coefficients $\bvec{y}$, such that the intensity of the map projected onto the sky of the observer is
$$I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{R} \, \bvec{y}.$$\\
\subsection{Rotation light curve}
Using this expression, the integrated intensity of the rotated map is
$$F = \oiint I(x,y)\,dS = \oiint \bvec{\tilde y}^\mathsf{T}\bvec{R}\bvec{y}\,dS$$
In order to simplify this integral, \citealt{starry} introduce a change of basis from the spherical harmonics basis $\bvec{\tilde y}$ to a simple polynomial basis $\bvec{\tilde p}$ characterized by the change of basis matrix $\bvec{A_1}$. In this new basis, $\bvec{y}$ can be expressed as
$$\bvec{p} = \bvec{A_1}\bvec{y}.$$
Hence, the integrated intensity becomes
$$F = \bvec{r^\mathsf{T}}\bvec{A_1}\bvec{R}\bvec{y},$$
with
$$\bvec{r^\mathsf{T}} \equiv \oiint \mathbf{\tilde{p}}^\top\,dS.$$
As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}$ can be computed analytically and only once for a given map degree $l_{max}$ (see \citealt[Eq. 20]{starry}).\\\\
\subsection{Occultation light curve}
The same formalism can be used to compute the integrated intensity of a surface map occulted by another spherical body. In this case, the intensity of the map must be integrated only over the unocculted surface $S$ of the disk, such that 
$$ F = \oiint_S I(x,y)\,dS = \left(\oiint_S \bvec{\tilde p^\mathsf{T}}\,dS\right)\bvec{A_1}\bvec{R}\bvec{y}.$$
To simplify this expression, \citealt{pal2012} note that any two-dimensional integral of the unocculted disk can be reduced to a one-dimensional integral over the contour $C$ of the occulting body using Green's theorem\footnote{Also known as Stokes' theorem in three-dimensional vector calculus.}. This can be done if the integrand of the integral over S can be described by a function $f$ such that
$$f(x, y) = \frac{\delta G_{y}}{\delta x}-\frac{\delta G_{x}}{\delta y},$$
where $G$ is a vector field defined and differentiable over $S$.\\\\
By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \citealt{starry} transforms $\bvec{\tilde p}$ to a basis $\bvec{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
$$s_n \equiv \oiint_S \tilde{g_n}(x, y)\,dS = \oint_C \bvec{G}_n(x, y)\,d\bvec{r},$$
with $\bvec{G}_n$ given in \citealt[Equation 34]{starry}. To simplify the bounds of this integral, the surface map can be aligned with the occulting body through a final rotation matrix $\bvec{R'}.$\\\\
Using these expressions, the integrated intensity of the occulted map follows the idiomatic linear expression 
$$F = \bvec{s^\mathsf{T}}\bvec{A}\,\bvec{R'}\,\bvec{R}\,\bvec{y},$$
with $\bvec{A} = \bvec{A_2A_1}$ and the integrals $s_n$ arranged in the \textit{solution vector} $\bvec{s^\mathsf{T}}$.\\\\

\subsection{Limb-darkened surfaces}

>>>>

The surface of a spherical body can be naturally described by spherical harmonics: a set of special functions that can be used as orthonormal basis to represent any spherical surface map. If $\bvec{\tilde y}$ denotes the spherical harmonics basis of degree $l_{max}$ (defined in \citealt[section X]{starry}) and $\bvec{y}$ a vector that contains the coefficients describing the surface in this basis, the intensity $I$ of the map projected onto the sky-projected $(x, y)$ plane of the observer is
$$I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{y}.$$
In order to orient this surface, for example with a given inclination, obliquity and phase around its rotation axis, we define a rotation matrix $\bvec{R}$ that acts on the spherical harmonics coefficients $\bvec{y}$, such that the intensity of the map projected onto the sky of the observer is
$$I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{R} \, \bvec{y}.$$\\
\subsection{Rotation light curve}
Using this expression, the integrated intensity of the rotated map is
$$F = \oiint I(x,y)\,dS = \oiint \bvec{\tilde y}^\mathsf{T}\bvec{R}\bvec{y}\,dS$$
In order to simplify this integral, \citealt{starry} introduce a change of basis from the spherical harmonics basis $\bvec{\tilde y}$ to a simple polynomial basis $\bvec{\tilde p}$ characterized by the change of basis matrix $\bvec{A_1}$. In this new basis, $\bvec{y}$ can be expressed as
$$\bvec{p} = \bvec{A_1}\bvec{y}.$$
Hence, the integrated intensity becomes
$$F = \bvec{r^\mathsf{T}}\bvec{A_1}\bvec{R}\bvec{y},$$
with
$$\bvec{r^\mathsf{T}} \equiv \oiint \mathbf{\tilde{p}}^\top\,dS.$$
As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}$ can be computed analytically and only once for a given map degree $l_{max}$ (see \citealt[Eq. 20]{starry}).\\\\
\subsection{Occultation light curve}
The same formalism can be used to compute the integrated intensity of a surface map occulted by another spherical body. In this case, the intensity of the map must be integrated only over the unocculted surface $S$ of the disk, such that 
$$ F = \oiint_S I(x,y)\,dS = \left(\oiint_S \bvec{\tilde p^\mathsf{T}}\,dS\right)\bvec{A_1}\bvec{R}\bvec{y}.$$
To simplify this expression, \citealt{pal2012} note that any two-dimensional integral of the unocculted disk can be reduced to a one-dimensional integral over the contour $C$ of the occulting body using Green's theorem\footnote{Also known as Stokes' theorem in three-dimensional vector calculus.}. This can be done if the integrand of the integral over S can be described by a function $f$ such that
$$f(x, y) = \frac{\delta G_{y}}{\delta x}-\frac{\delta G_{x}}{\delta y},$$
where $G$ is a vector field defined and differentiable over $S$.\\\\
By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \citealt{starry} transforms $\bvec{\tilde p}$ to the \textit{Green's basis} $\bvec{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
$$s_n \equiv \oiint_S \tilde{g_n}(x, y)\,dS = \oint_C \bvec{G}_n(x, y)\,d\bvec{r},$$
with $\bvec{G}_n$ given in \citealt[Equation 34]{starry}. To simplify the bounds of this integral, the surface map can be aligned with the occulting body through a final rotation matrix $\bvec{R'}.$\\\\
Using these expressions, the integrated intensity of the occulted map follows the idiomatic linear expression 
$$F = \bvec{s^\mathsf{T}}\bvec{A}\,\bvec{R'}\,\bvec{R}\,\bvec{y},$$
with $\bvec{A} = \bvec{A_2A_1}$ and the integrals $s_n$ arranged in the \textit{solution vector} $\bvec{s^\mathsf{T}}$.\\\\

\subsection{Limb-darkened surfaces}
We notice that a simpler situation is encountered when the surface of the spherical body can be described by a radial intensity profile $I(r)$, such as a limb-darkened star. In this case, the vector of spherical harmonics coefficients $\bvec{y}$ is very sparse (because all $m\neq0$ components are zero) and the surface map doesn't need to be rotated. Contemporary to \cite{starry}, \cite{Agol2020} explore this case and provide simple expressions for the integrated intensity of a star whose surface can be described by a polynomial limb-darkening law occulted by an opaque body. In this case:
\begin{enumerate}
    \item The surface is simply described by $\bvec{u}$, a vector of polynomial limb-darkening coefficients.
    \item The Green's basis takes a simpler form (see \citealt[Eq. X]{Agol2020}).
    \item Matrices involved in the computation of the integrated flux have $l_{max}$ columns, versus $(l_{max}+1)^2$ in the general case, reducing drastically the computational cost of the model.
\end{enumerate}
For these reasons, and because of its wide application, our implementation treats radially symmetric maps as a separate case.\\\\
